---
title: "PaperPlotR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PaperPlotR}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[UTF-8]{inputenc}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(ggplot2)
library(PaperPlotR)
```

`PaperPlotR` is designed for scientific figure workflows that rely on
`ggplot2`, but need additional structure around styling, figure sizing,
semantic colors, layout, and export.

The package aims to standardize repeated plotting decisions without hiding the
underlying `ggplot2` workflow.

## Inspect Available Presets

```{r}
available_fig_specs()
available_journal_presets()
available_output_presets()
available_group_dictionaries()
```

## Figure Size Helpers

`fig_specs` stores the built-in panel definitions used by `get_fig_size_cm()`
and `panel_size()`.

```{r}
fig_specs
get_fig_size_cm("4.9x2", ncol = 3, nrow = 1)
journal_preset("cell")
recommend_panel_spec(n_panels = 4)
```

For dense figures, `recommend_panel_spec()` can suggest a starting layout and
indicate when splitting a figure is likely to be clearer.

```{r}
recommend_panel_spec(
  n_panels = 8,
  heatmap = TRUE,
  rotated_x_labels = TRUE,
  complex_legend = TRUE
)
```

You can extend the defaults at runtime with custom registrations.

```{r}
register_fig_spec(
  name = "6x4",
  panel_w_cm = 6,
  panel_h_cm = 4,
  layout_hint = "Wide single-panel figures"
)

register_palette(
  name = "study_palette",
  values = c("#1B3A57", "#2E6F95", "#58A4B0")
)

tail(available_fig_specs(), 1)
tail(available_lab_palettes(), 1)
```

## Theme and Color Scales

```{r}
p <- ggplot(mtcars, aes(wt, mpg, colour = factor(cyl))) +
  geom_point(size = 2.2) +
  scale_color_lab() +
  theme_lab()

p
```

Use semantic group mappings when category names should keep consistent colors
across plots.

```{r}
group_colors("default")[1:4]
```

## Layout and Export

```{r}
p1 <- ggplot(mtcars, aes(wt, mpg)) + geom_point() + theme_lab()
p2 <- ggplot(mtcars, aes(factor(cyl), mpg)) + geom_boxplot() + theme_lab()

layout_lab(p1, p2, ncol = 2)
```

Use `save_lab()` when you want figure size to follow a named panel spec plus a
journal preset.

```r
save_lab(
  plot = p1,
  filename = "figures/scatter_cell.png",
  spec = "4.9x4.9",
  journal = "cell"
)
```

Use `save_lab_plot()` when you prefer named output presets or direct width and
height control.

## Comparison Plots

`PaperPlotR` also provides lightweight wrappers for common grouped comparison
plots.

```{r}
df <- data.frame(
  condition = rep(c("Control", "Treatment"), each = 12),
  value = c(rnorm(12, 5, 0.4), rnorm(12, 6.2, 0.5))
)

plot_box_paper(
  df,
  condition,
  value,
  dictionary = "default",
  show_signif = TRUE,
  comparisons = list(c("Control", "Treatment"))
)
```

If you prefer lower-level control, the same workflow can be built from
`layer_points_paper()`, `layer_summary_paper()`, and `layer_signif_paper()`.
